### 2024-10-10 Article/main.tex
osservazioni

\begin{definition}[Category]\label{def:category}
    A \emph{category} $\cat{C}$ comprises:
    \begin{enumerate}
        \item A collection of \emph{objects} $\Ob(\cat{C})$;
        \item A collection of \emph{arrows} (or \emph{morphisms}) $\Hom(\cat{C})$, often called \emph{homset}.
    \end{enumerate}
    For each morphism $f \in \Hom(\cat{C})$, there are two operators $dom$ and $cod$ that map every morphism to two objects, respectively, its \emph{domain} and its \emph{codomain}. In case $dom\ f = A$ and $cod\ f = B$, we will write $f: A \rightarrow B$. The collection of morphisms from an object $A$ to an object $B$ is denoted as $\cat{C}(A, B)$.
    An operator $\circ$ of \emph{composition} maps every couple of morphisms $f$, $g$ with $cod\ f = dom \ g$ (in this case $f$ and $g$ are said to be composable) to a morphism $g \circ f : dom\ f \rightarrow cod \ g$. The composition operator is associative, i.e., for each composable arrows $f$, $g$ and $h$, it holds that
    $$
        h \circ (g \circ f) = (h\circ g) \circ f
    $$
    For each object $A$, an \emph{identity} morphism $id_A : A \rightarrow A$ (or, when it is clear from the context, just denoted $A$)  such that, for each $f: A \rightarrow B$:
    \[
        id_B \circ f = f = f \circ id_A 
    \]
\end{definition}

@@

\begin{example}
    It is easy to see that taking sets as objects and total functions as arrows, we obtain a category. In fact, given two functions $f: A \rightarrow B$ and $g: B \rightarrow C$, it is possible to compose them obtaining an arrow $g \circ f : A \rightarrow C$, and the composition is associative. For each set $A$ there exists an identity function $id_A: A \rightarrow A$ such that $id_A(a) = a$ for each $a\in A$.
    This category is denoted as $\Set$.
\end{example}

@@

\begin{definition}[Subobjects]
    Let $C$ be an object in a category $\cat C$. Then, if $m: A \rightarrow C$ is mono, $(A, m)$ is said to be a \emph{subobject} of $C$. Factorization of morphisms induces a preorder on subobjects of an object. $(A, m) \leq (B, n)$ whenever there exists a morphism $f : A\rightarrow B$ such that $m = n \circ f$.
    \[
        \begin{tikzcd}[row sep=9]
            A \arrow[dr, "m"] \arrow[dd, dashed, "f"swap] & \\
            & B \\
            C \arrow[ur, "n"swap] &
        \end{tikzcd}
    \]
    Two subobject $(A, m)$ and $(B, n)$ can are said to be \emph{equivalent subibjects}, written $(A, m) \approx (B, n)$ if $(A, m)\leq (B, n)$ and $(B, n) \leq (A, m)$.
\end{definition}

@@

\begin{prop}\label{prop:epi_mono_prop}
    The following statements hold for every pair of composable arrows $f$ and $g$ for any category $\cat C$:
    \begin{enumerate}
        \item if both $f$ and $g$ are mono, then $g \circ f$ is mono;
        \item if $g \circ f$ is mono, then $f$ is mono;
        \item if both $f$ and $g$ are epi, then $g \circ f$ is epi;
        \item if $g \circ f$ is epi, then $g$ is epi.
    \end{enumerate}
\end{prop}

@@

\todo{E cambierei anche l'intro a questa sezione DC}

>>>>

\begin{definition}[Category]\label{def:category}
    A \emph{category} $\cat{C}$ comprises:
    \begin{enumerate}
        \item A collection of \emph{objects} $\Ob(\cat{C})$;
        \item A collection of \emph{arrows} (or \emph{morphisms}) $\Hom(\cat{C})$, often called \emph{homset}.
    \end{enumerate}
   \todo{Non sono d'accordo con quello che è scritto qui. Non è vero che per ogni morfismo esistono due operatori dom e cod. Semmai esistono due operatori che ad ogni morfismo associano un dominio ed un codominio} For each morphism $f \in \Hom(\cat{C})$, there are two operators $dom$ and $cod$ that map every morphism to two objects, respectively, its \emph{domain} and its \emph{codomain}. In case $dom\ f = A$ and $cod\ f = B$, we will write $f: A \rightarrow B$. The collection of morphisms from an object $A$ to an object $B$ is denoted as $\cat{C}(A, B)$.
    An operator $\circ$ of \emph{composition} maps every couple of morphisms $f$, $g$ with $cod\ f = dom \ g$ (in this case $f$ and $g$ are said to be composable) to a morphism $g \circ f : dom\ f \rightarrow cod \ g$. The composition operator is associative, i.e., for each composable arrows $f$, $g$ and $h$, it holds that
    $$
        h \circ (g \circ f) = (h\circ g) \circ f
    $$
    For each object $A$, an \emph{identity} morphism $id_A : A \rightarrow A$ (or, when it is clear from the context, just denoted $A$)  such that, for each $f: A \rightarrow B$:
    \[
        id_B \circ f = f = f \circ id_A 
    \]
\end{definition}

@@

\begin{example}
    It is easy to see that taking sets as objects and total \todo{Ha senso mettere totale qui?}functions as arrows, we obtain a category. In fact, given two functions $f: A \rightarrow B$ and $g: B \rightarrow C$, it is possible to compose them obtaining an arrow $g \circ f : A \rightarrow C$, and the composition is associative. For each set $A$ there exists an identity function $id_A: A \rightarrow A$ such that $id_A(a) = a$ for each $a\in A$.
    This category is denoted as $\Set$.
\end{example}

@@

\begin{definition}[Subobjects]
    Let $C$ be an object in a category $\cat C$. Then, if $m: A \rightarrow C$ is mono\todo{sinceramente non trovo molto sensata la scelta di chiamare sottoggetto la coppia $(A, m)$. Lo standard, per quanto ne so, è usare sottoggetto per una classe di equivalenza di mono o, al massimo, per i mono}, $(A, m)$ is said to be a \emph{subobject} of $C$. Factorization of morphisms induces a preorder on subobjects of an object. $(A, m) \leq (B, n)$ whenever there exists a morphism $f : A\rightarrow B$ such that $m = n \circ f$.
    \[
        \begin{tikzcd}[row sep=9]
            A \arrow[dr, "m"] \arrow[dd, dashed, "f"swap] & \\
            & B \\
            C \arrow[ur, "n"swap] &
        \end{tikzcd}
    \]
    Two subobject $(A, m)$ and $(B, n)$ can are said to be \emph{equivalent subobjects}, written $(A, m) \approx (B, n)$ if $(A, m)\leq (B, n)$ and $(B, n) \leq (A, m)$. 
\end{definition}

@@

\begin{prop}\label{prop:epi_mono_prop}
    The following statements hold for every pair of composable arrows $f$ and $g$ for any category $\cat C$:
    \begin{enumerate}
        \item if both $f$ and $g$ are mono, then $g \circ f$ is mono;
        \item if $g \circ f$ is mono, then $f$ is mono;
        \item if both $f$ and $g$ are epi, then $g \circ f$ is epi;
        \item if $g \circ f$ is epi, then $g$ is epi.
    \end{enumerate}
\end{prop}
\todo{val la pena osservare la relazione di equivalenza tra sottoggetti corrisponde ad avere un isomorfismo tra i domini}

@@

\todo{E cambierei anche l'intro a questa sezione DC}
\todo{la 1.1 è un po' povera come sezione, se sposti questa sezione sui funtori prima, poi puoi arricchire la 1.1 con la sottosezione sulle comma}
